Can We Crunch the Social Graph on the Cheap?

### 2011-01-25 doc/hotpar2011/softxmt-hotpar2011.tex
modified:   softxmt-hotpar2011.tex
system overview and intro; additional authors

\documentclass{acm_proc_article-sp}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{alltt}
\usepackage{comment}
\begin{document}

@@

\numberofauthors{7}
\author{
Jacob Nelson, Brandon Myers, Andrew Hunter, Dan Grossman, Luis Ceze\\
University of Washington\\
\email{\{nelson, bdmyers, ahh, djg, luisceze\}@cs.washington.edu}\\ \\
Simon Kahan \\ 
Pacific Northwest National Lab\\
\email{\{skahan\}@cs.washington.edu}
}

@@

We want the best of both worlds. Our goal is to build a system that
has good performance on these low-locality graph codes but is
implemented with commodity processors. The core of our approach is a
lightweight coroutine library designed to overlap long-latency memory
prefetches with other contexts' computation.

In this paper, we describe our single-node runtime for latency
tolerance, along with our plans for scaling to multiple nodes.  We
evaluate the performance of our runtime using a set of list traversal
benchmarks designed to model certain worst-case behaviors of these
large graph problems.

The rest of the paper is organized as follows. We present related work
in section~\ref{sec:related}. We describe the programming model we are
implementing in section~\ref{sec:model}, and describe our approach for
implementing this model in section~\ref{sec:approach}. We evaluate our
latency-tolerant runtime in section~\ref{sec:evaluation} and conclude
in section~\ref{sec:conclusion}.

@@

Our programming model partitions the address space into a shared
global shared address space and per node private address spaces.
Locality may be exploited directly by the programmer in the node
private address spaces, just as in a conventional cache-coherent
multiprocessor node.  In contrast, the global space is explicitly
uncached.  There, every processor can access any word of memory
directly via its unique global address.

Whereas in the PGAS model coherency of global copies is implictly
maintained across the system, we forego caching of globally shared
data.  A node can exploit locality within a chunk of data stored in
the global space only by first copying it into the node's local space.
In this way, a node can access data at random across the system while
incurring no hardware coherence overhead: consistency is enforced by
our mechanisms for explicit global synchronization.

The model promises efficiency subject to the presence of sufficient
concurrency.  Exactly how that concurrency is expressed by the
programmer is language dependent.  Our programming model serves to
augment threaded programming languages so that they become latency
tolerant on large address spaces.  Software multithreading is used to
provide latency tolerance: a computation that is about to execute a
long latency operation such as a remote memory reference initiates the
operation and then yields to the scheduler, which quickly resumes
execution of another computation.  Concurrency is also used -- via
aggregation of requests to a common node destination -- to make more
efficient use of network bandwidth and in some cases reduce bandwidth
demands.

Synchronization on locations in the node private address space is the
same as on conventional systems.  Synchronization on global addresses
works differently: we provide atomic operations such as {\tt
  int\_fetch\_add} as well as operations using full-bits, as on the
Cray XMT.  As is true for other long latency operations,
synchronization latency is tolerated by yielding to the scheduler.
Even non-blocking global atomics yield: atomic operations executed as
x86 instructions sabbotage concurrency by forcing all outstanding
memory references to complete before new ones can be issued.  By
yielding, the core can switch to another computation while the
synchronization is accomplished outside of the pipeline.  In addition,
the implementation can exploit aggregation when concurrency is
abundant to increase the throughput of global synchronization.

Thus, the programming model provides the programmer with options: express
locality within the private space, and it will be exploited to reduce
latency and bandwidth; express concurrency, and it will be used toward
tolerating latency and reducing the cost of bandwidth.

\begin {comment}
\todo{Change to discuss API:

@@

\end{comment}

>>>>

\documentclass{acm_proc_article-sp}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{alltt}
\begin{document}

@@

\numberofauthors{7}
\author{
Jacob Nelson, Brandon Myers, Andrew Hunter, Dan Grossman, Mark Oskin,
Carl Ebeling, Luis Ceze\\
University of Washington\\
\email{\{nelson, bdmyers, ahh, djg, oskin, ebelin, luisceze\}@cs.washington.edu}\\ \\
Simon Kahan \\ 
Pacific Northwest National Lab\\
\email{\{skahan\}@cs.washington.edu}
Preston Briggs\\ 
Cray Inc.\\
\email{\{preston.briggs\}@gmail.com}
}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=0.9\textwidth]{figures/system-overview.pdf}
	\end{center}
	\caption{Overview of our proposed system design. The base
          system is a cluster of commodity nodes and interconnect. We
          add a latency-tolerant runtime and a global memory manager
          to obtain efficient access to the global address space. Only
          the shaded components may require hardware design.}
	\label{fig:system-overview}
\end{figure*}

We want the best of both worlds. Our goal is to build a system that
has good performance on these low-locality graph codes but is
implemented with cheap commodity processors.

Figure~\ref{fig:system-overview} is a schematic of such a
system. It is composed of multiple nodes built with commodity
processors, communicating over an Infiniband network. We add two
components: a runtime system, responsible for executing the user's threads,
and a global memory manager, responsible for facilitating memory
operations to the non-coherent global memory shared across the nodes.
This memory manager is the only component that might require hardware,
but we are exploring a mix of hardware and software implementations.

This paper explores the feasibility of our multi-node goal using a
single-node implementation of our lightweight threading runtime. We
use prefetch instructions together with lightweight threads to
tolerate the latency of DRAM access on the local node node. We use
pointer chasing benchmarks to verify our runtime's overhead is
acceptable. We also model the effects of pointer chasing on a remote
node.

The rest of the paper is organized as follows. We describe our
proposal from a programmer's viewpoint in section~\ref{sec:model}. We
describe the implementation of our latency-tolerant runtime and plans
for extending to multiple nodes in section~\ref{sec:approach}. We
evaluate our runtime in section~\ref{sec:evaluation}.  We present
related work in section~\ref{sec:related} and conclude in
section~\ref{sec:conclusion}.

@@

\todo{Change to discuss API:

@@
